Transcriptomic Analysis of Expressions of Biomarkers and Their Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-06-27 Cover_letter_JPM.tex
Updates from Overleaf

[Discussion]
[more in the discussion section]
1) 
Statistical significance (\textit{P} value) is affected by both sample size, error, effect size (substantive significance)\cite{Sullivan2012}\cite{Thiese2016}, and cutoff. 
The effect size is the magnitude of the difference (or relationship) among comparing groups.
%A \textit{P} value is a tool to help interpret findings from biomedical research.
%In reporting and interpreting studies, both 
%the substantive significance (effect size) and statistical significance (P value) are essential results to be reported.
While statistical significance can inform the researcher whether an effect exists, 
the \textit{P} value will not directly tell the effect size.
%The effect size is the main finding of a quantitative study.
%Statistical significance depends upon both sample size and effect size. 
The effect size is independent of the sample size\cite{Sullivan2012}.

@@

As skewed cutoff that split 425 vs 75 between the two groups, it will also get statistical significance by increasing effect size intentionally.
% https://stats.stackexchange.com/questions/162359/how-can-you-have-p-0-0001-with-a-sample-size-of-89-and-a-control-group-of-52
%As you can see, even with only 12 events (fustat = 1) that split 10 vs 2 between the two dummy groups you get a log-rank p-value < 0.0001.
%effect size, sample size, cutoff 影響 p value!

@@

2)
cutoff at optimal by sliding-window or just median:
when \textit{P} value $<0.05$, the gene number: 9416 , 534 %差別很大，張大網
run code: if median cut for TCGA KM Cox (by pvalueTex or *** RLassoCox Bayes code) 9416 -> ???
discussion about sliding-window pvalueTex to find the more potential candidate, while small effect size could be double-checked by GSE65858.
(  )2-3 put in supplementary figure Rplot_pvaluePlot_NDFIP1.pdf
%列入 workflow in materials and methods;

***NDFIP1
a special case of NDFIP1 gene; % 給他們一個機會
(ok) Answer 2-1 %加入 discussion
( )  Answer 2-3 supplementary  %圖就好 figure Rplot_pvaluePlot_NDFIP1.pdf

For example, a gene NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile cutoff, could achieve a \textit{P} value of \num{1e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-corrected \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233}.
However, it fails as the candidate by inadequate effect size (HR = ??).
% (ok) Answer 2-1 加入 discussion
According to the results of sliding-window cutoff finding, the preliminary biomarkers have FDR-corrected \textit{P} values from \num{1e-4} to 0.04 which have been grouped by cutoffs between 50\% and 70\% quantile (Please see Figure \ref{fig:figure4}; it is also shown in manuscript as updated "Figure 4": top 3 candidate biomarkers, CAMK2N1, IL19, and FCGBP).
Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in 
supplementary figure S1). After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233}.
%Rplot_pvaluePlot_NDFIP1.pdf
%  to find cases_OS outside of range 180-220 (== cutoff not at median)
%# pvalue plot as 倒V字形 "^" or W p_OS pvalues plots (with local minials)
% # 99 genes have one p_OS >= 0.05, in OS_pvalueTex005
%# 3641 genes have one p_OS >= 0.02, in OS_pvalueTex002
%# special case: HNSCC_survivalAnalysis_marginS_NDFIP1.xlsx
%# NDFIP1 (within our 20 preliminary candidates), 
In those \textit{P}-value plots, we could see that the sliding-window selection of cutoff finding might collect more potential candidates (\textit{P} values far less than $0.001$) for further Cox's modeling.
%Some useful biomarkers might be kick-out prematurely in "median cutoff" of survival analysis.
Indeed, it needs to be validated by the other independent cohort (e.x. GSE65858).

@@

% dissection
%ZSWIM2\_archive.Rda % 其實最困難的在於 unknown error (ZSWIM2) 
Third, we have analyzed the error log of the cutoff finding and Cox modeling, 
the results shows that several situations could halt the program running.
It includes:
\begin{outline}
\1 32.2\% event has one group issue in confusion matrix of Chi-square test in Cox regression
\1 21.05\% errors occur by one group issue in log-rank test (survdiff function in R package "survival") in Kaplan-Meier estimate
\1 0.78\% has unknown reason (so those 159 genes has been excluded in our workflow)
\end{outline}
, which could not be detected prior to the program running.
It might be due to heterogeneity of 20500 RNA-Seq expression data.
%# deal with return(1), return(2) and return(3) errors, try to solve it
%32.2\% error01, ZSWIM2 skip from contingencyTCGA(), it is one group issue in Melt() function.
%21.05\% errors in  one group in survdiff on KM cutofFinder_func.R.
%0.78\%(159 genes) for unknown reason
%# _marginS_
%n_percent_Bonferroni <- ZSWIM2_2$Freq[1]/sum(ZSWIM2_2$Freq) # 0.4593171 -> 0.4581951
%error01_sample <- which(ZSWIM2$X3==1) #  32.2\% error01: "ZSWIM2 skip from contingencyTCGA()"): one group issue in Melt()
%# => *** re-run
%error02_sample <- which(ZSWIM2$X3==2) # 14.2\% error02: There has only one group in survdiff.
%error03_sample <- which(ZSWIM2$X3==3) # 6.85\% error03: There has only one group in survdiff on KM cutofFinder_func.R.
%#=> Test for difference (log-rank test is multiple Chi-square alone survival time) with P-value; or Test for different KM curve between two groups (seperated by PMM1_median 0 vs 1)
% error05_sample <- which(ZSWIM2$X3==5) # 0.78\%(159 genes) error05: for why?

>>>>

[Discussion]
[more in the discussion section]
1) % (ok)
Statistical significance (\textit{P} value) is affected by both sample size, error, effect size (substantive significance)\cite{Sullivan2012}\cite{Thiese2016}, and cutoff. 
The effect size is the magnitude of the difference (or relationship) among comparing groups.
%A \textit{P} value is a tool to help interpret findings from biomedical research.
%In reporting and interpreting studies, both 
%the substantive significance (effect size) and statistical significance (P value) are essential results to be reported.
While statistical significance can inform the researcher whether an effect exists, 
the \textit{P} value will not directly tell the effect size.
%The effect size is the main finding of a quantitative study.
%Statistical significance depends upon both sample size and effect size. 
The effect size is independent of the sample size\cite{Sullivan2012}.

@@

If a skewed cutoff that split, for example, 425 vs 75 between the two groups, it will also get statistical significance by increasing effect size artificially.
% https://stats.stackexchange.com/questions/162359/how-can-you-have-p-0-0001-with-a-sample-size-of-89-and-a-control-group-of-52
%As you can see, even with only 12 events (fustat = 1) that split 10 vs 2 between the two dummy groups you get a log-rank p-value < 0.0001.
%effect size, sample size, cutoff 影響 p value!

@@

2) in TCGA HNSCC cohort, 
cutoff at optimal by sliding-window or just median:
when unadjusted \textit{P} value $<0.05$, the gene number: 9416 to 3118, 534 to 209.
Result by median cutoff: total 20080 genes 
%3118 are raw KM P value < 0.05; 
%209 genes are FDR KM P value <0.05
*** IL19 is excluded by this way :-) (special S-curve P value plot)
%差別很大，張大網
%([2021/06/27] run code: if median cut for TCGA KM Cox by *** pvalueTex code)
discussion about sliding-window pvalueTex to find the more potential candidate, while small effect size could be double-checked by GSE65858.
(  )2-3 put in supplementary figure %Rplot_pvaluePlot_NDFIP1.pdf
%列入 workflow in materials and methods;

% ****** IL19 必須要靠 pvalueTex 才能發現
\textit{P} value plot with a acute S-curve bending on the median zone, it lets the FDR-KM \textit{P} value has large difference between median cutoff (\num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.1154315}) and optimal cutoff (\num{6.54e-6}).
A sliding-window selection of cutoffs could boost their statistic significance to pass the correction by even FDR or Bonferroni methods.
%> gene_labeled_golden[!gene_labeled_golden %in% fdr100_KMsurvival$Gene.Symbol[fdr100_KMsurvival$FDR.Pvalue<0.05]]
[1] "IL19"
***NDFIP1
the special cases of gene, such as NDFIP1, DKC1, PNMA5, and NPB % ; % 給他們一個機會
(ok) Answer 2-1 %加入 discussion
( )  Answer 2-3 supplementary  %圖就好 figure Rplot_pvaluePlot_NDFIP1.pdf

For example, a gene NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile cutoff, could achieve a \textit{P} value of \num{1e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-corrected \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233}.
However, it fails as the candidate by less effect size (HR = 1.33 in GSE65858 cohort, comparing with HR = 2.027 in TCGA cohort).
% (ok) Answer 2-1 加入 discussion
According to the results of sliding-window cutoff finding, the preliminary biomarkers have FDR-corrected \textit{P} values from \num{1e-4} to 0.04 which have been grouped by cutoffs between 50\% and 70\% quantile (Please see Figure \ref{fig:figure4}; it is also shown in manuscript as updated "Figure 4": top 3 candidate biomarkers, CAMK2N1, IL19, and FCGBP).
Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in 
supplementary figure S1). After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233}.
%Rplot_pvaluePlot_NDFIP1.pdf
%  to find cases_OS outside of range 180-220 (== cutoff not at median)
%# pvalue plot as 倒V字形 "^" or W p_OS pvalues plots (with local minials)
% # 99 genes have one p_OS >= 0.05, in OS_pvalueTex005
%# 3641 genes have one p_OS >= 0.02, in OS_pvalueTex002
%# special case: HNSCC_survivalAnalysis_marginS_NDFIP1.xlsx
%# NDFIP1 (within our 20 preliminary candidates), 
In those \textit{P}-value plots, we could see that the sliding-window selection of cutoff finding might collect more potential candidates (\textit{P} values far less than $0.001$) for further Cox's modeling.
%Some useful biomarkers might be kick-out prematurely in "median cutoff" of survival analysis.
Indeed, it needs to be validated by the other independent cohort (e.x. GSE65858).

@@

% dissection
%ZSWIM2\_archive.Rda % 其實最困難的在於 unknown error (ZSWIM2) 
Third, we have analyzed the error log of the cutoff finding and Cox modeling, 
the results shows that several technique problems could halt the program running.
It includes:
\begin{outline}

\1 32.2\% event has "one group" issue in confusion matrix of Chi-square test in Cox regression
\1 21.05\% error occurs by "one group" issue in log-rank test (survdiff or survdiff.fit function in R package "survival") in Kaplan-Meier estimate.
%1) ok [1] "[ 4 / 20083 ] Generating median-cut KM plot of A1BG"
%Error in survdiff.fit(y, groups, strata.keep, rho) : at A1CF
% error at 16115 gene ""; why? 518 values all are -0.04203 at T
%  There is only 1 group 
%This is because of skewed distribution of the expression value. 
% 67/518 =  13% cases >1, while 450 (87%) cases = -0.30949; so median is -0.30949; 
% all cases belong to >= median => 改為 > median 即可
% 2) % survfit 16114 has data set has no non-missing observations: NA's 113/518 cases
% median = NaN :-); 
% solved by cutoff <- median(gset_OS_mRNA2[, i], na.rm = TRUE)
% and NaN imputation as median
\1 0.78\% has unknown reason (so those 159 genes has been excluded in our workflow)
\end{outline}
, which could not be detected prior to the program running.
It might be due to skewed distribution of the expression value in 20500 RNA-Seq  data. % heterogeneity
%# deal with return(1), return(2) and return(3) errors, try to solve it
%32.2\% error01, ZSWIM2 skip from contingencyTCGA(), it is one group issue in Melt() function.
%21.05\% errors in  one group in survdiff on KM cutofFinder_func.R.
%0.78\%(159 genes) for unknown reason
%# _marginS_
%n_percent_Bonferroni <- ZSWIM2_2$Freq[1]/sum(ZSWIM2_2$Freq) # 0.4593171 -> 0.4581951
%error01_sample <- which(ZSWIM2$X3==1) #  32.2\% error01: "ZSWIM2 skip from contingencyTCGA()"): one group issue in Melt()
%# => *** re-run
%error02_sample <- which(ZSWIM2$X3==2) # 14.2\% error02: There has only one group in survdiff.
%error03_sample <- which(ZSWIM2$X3==3) # 6.85\% error03: There has only one group in survdiff on KM cutofFinder_func.R.
%#=> Test for difference (log-rank test is multiple Chi-square alone survival time) with P-value; or Test for different KM curve between two groups (seperated by PMM1_median 0 vs 1)
% error05_sample <- which(ZSWIM2$X3==5) # 0.78\%(159 genes) error05: for why?
